---
title: "Leukemia patients - Survival Analysis"
output:
  word_document: default
  html_document:
    pdf_print: 
     
  pdf_document: default
---

# Description:

This report analyzes the effect of a treatment on 42 children with acute leukemia . Patients were observed until they relapsed ,dropped out ,or the study ended.

# Methodology:

This report uses non-parametric techniques to observe difference between the placebo and the treatment group. A Kaplan-Meier Estimates is computed for each group in order to observe the survival function of each independently . A log rank test is then performed to test whether both groups are significantly different or not.

## Variable Descriptions

-   Time : *Represents the time to relapse/censor for each patient*

-   Treatment : *Represents whether subject received a treatment or not (treated=1, placebo=0 )*

-   Failure : *Represents whether subject relapsed or censored (relapsed =1 , censored = 0 )*

## Imports

```{r message=FALSE, warning=FALSE, results='hide'}
library(readxl)
library(ggplot2)
library(dplyr)
library(epiDisplay)
library(epicalc)
library(survival)
library(survminer)

```

```{r}
df <- read_excel("leukemia.xlsx") # total ob.servations 42
df <- df[-c(4:7)] # dropping description columns
```

### Wrangling

```{r}
groupss <- factor(df$Treatment,labels=c("Placebo","Treatment"))
event <- factor(df$Failure,labels=c("Censored","Died"))
df1<- cbind(df,event)
ID<- c(1:42)
df1<- cbind(df,ID)
df1$Treatment <- groupss
```

#### Follow-Up graph

```{r warning=FALSE}

ggplot(df1, 
       aes(x=df1$Time, 
           y=df1$ID)) +
  geom_segment(aes(x = 1, 
                   xend = df1$Time, 
                   y = df1$ID, 
                   yend = df1$ID),
               color = "lightgrey") +
  labs (x = "Time to Death/ Censored (years)",
        y = "Patient ID",
        title = "Time to Death",
        subtitle = "Follow-up time to 42 patients") +
  theme_minimal()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  geom_point(aes(shape=event), size = 3) 
```

Patients 22 and above represents the placebo group

#### Failure vs Censoring 

```{r}
df3 <- df1
df3$Failure <- factor(df1$Failure,labels=c("Censored","Relapse"))
plot(df3$Failure)
```

## Kaplan-Meier Estimates / Treatment group

kaplan Meir refers to................

```{r}
df.treat <- df1[ which(df1$Treatment=='Treatment'),] # dividing into 2 groups 


KMTREAT <- survfit(Surv(df.treat$Time, df.treat$Failure) ~ 1,conf.type="log-log", data = df.treat)
summary(KMTREAT)

```

Comment

## Kaplan-Meier Estimates /placebo group

```{r}
df.placebo <- df1[ which(df1$Treatment=='Placebo'),]

KMPLACE <- survfit(Surv(df.placebo$Time, df.placebo$Failure) ~ 1,conf.type="log-log", data = df.placebo)
summary(KMPLACE)

```

## Survival-curves

```{r}
library(ggfortify)


combinedKM <- survfit(Surv(df1$Time, df1$Failure) ~ df1$Treatment, data = df1)

autoplot(combinedKM)
```

#### Comparing survival functions /Log Rank test

```{r}

ggsurvplot(combinedKM, data = df1, pval = TRUE, 
           xlab="Time to Relapse", surv.median.line = c("hv"))

survdiff(Surv(df1$Time, df1$Failure) ~ df1$Treatment, data = df1)
```
